### 2017-04-01 paper.tex
Keep adding to frame format

%TODO: mention one-way

As stated in chapter \ref{ch:literature-review}, the United Kingdom's Office
of Communications restrict what frequencies can be used for airborne
radio transmissions, and how much power the transmissions can be sent at.
Those restrictions constrain what existing protocols can be used to transmit
arbitrary data over radio. This disqualifies existing packet radio protocols
that would have fulfilled most of the requirements, in use in other countries
like AX.25.

@@

The AHABus radio protocol was designed to match the requirements for the radio
system as closely as possible. While the specification presented in this section
is the last version, which was implemented in the AHABus prototype, some parts
of the protocol evolved or had to be changed during the prototyping process.
% TODO: explain more? less?

@@

@@

\item[Frame Layer:] The frame layer relies on the Link Layer to transmit its
data units, and provides reliability and error correction to higher-level 
layers.

@@

\subsubsection{Link Layer}

@@

RTTY functions like UART, over radio. Instead of using logical electrical levels
(5V for logical high, 0V for logical low), it uses Audio Frequency Shift Keying
(\textit{AFSK}): two audio tones separated by a frequency shift act as two
logical levels. When nothing is being transmitted, the line is held high. The
start of a byte is signalled by the \textit{start bit} (the line goes low),
followed by the eight \textit{data bits}, and one or two \textit{stop bits} (the
line is held low). After a byte has been sent, the line is held high again until
the next byte, or the next byte starts immediately.

%TODO: replace mentions of UART with RS-232 // NOPE, RS232 diff. than TTL
%TODO: references for DominoEx and Olivia
Other existing protocols exist that can transmit data over radio audio signals,
like DominoEx or Olivia, sometimes with better transmission rates. RTTY was
preferred over them because the ESP8266 board used as OBC in the AHABus
prototype provides hardware support for UART, making it easier to implement and
reliable. Since the link layer is transparent to the frame and packet layers,
RTTY could be swapped for another protocol in the future with no impact on
those.

@@

\begin{figure}[H]
    \begin{bytefield}{16}
        \bitheader{0,7,8,15} \\
        \begin{rightwordgroup}{Header}
            \bitbox{8}{01011010} & \bitbox{8}{version} \\
            \bitbox{16}{sequence number}
        \end{rightwordgroup} \\
        \begin{leftwordgroup}{Payload}
            \wordbox[lrt]{1}{data} \\
            \skippedwords \\
            \wordbox[lrb]{1}{}
        \end{leftwordgroup} \\
        \begin{rightwordgroup}{FEC}
            \wordbox[lrt]{1}{checksum} \\
            \skippedwords \\
            \wordbox[lrb]{1}{}
        \end{rightwordgroup}
    \end{bytefield}
    \centering
    \caption{AHABus Frame Format}
\end{figure}

@@

\begin{figure}[H]
    \begin{bytefield}{32}
        \bitheader{0,7,8,15,16,23,24,31} \\
        \begin{rightwordgroup}{Header}
            \bitbox{8}{version} & \bitbox{8}{payload ID} &
            \bitbox{16}{packet length} \\
            \bitbox{32}{latitude} \\
            \bitbox{32}{longitude} \\
            \bitbox{16}{altitude} & \bitbox[lrt]{16}{}
        \end{rightwordgroup} \\
        \begin{leftwordgroup}{Payload}
            \wordbox[lrt]{1}{data} \\
            \skippedwords \\
            \wordbox[lrb]{1}{}
        \end{leftwordgroup}
    \end{bytefield}
    \centering
    \caption{AHABus Frame Format}
\end{figure}

@@

@@

\subsection{Reception Binary Stream Extraction}

@@

\newpage
\chapter{Results}
\label{ch:reults}

>>>>

As stated in chapter \ref{ch:literature-review}, the United Kingdom's Office
of Communications restrict what frequencies can be used for airborne
radio transmissions, and how much power the transmissions can be sent at.
Those restrictions constrain what existing protocols can be used to transmit
arbitrary data over radio. This disqualifies existing packet radio protocols
that would have fulfilled most of the requirements, in use in other countries
like AX.25. \todo{mention Tx only, no Rx}

@@

The AHABus radio protocol was designed to match the requirements for the radio
system as closely as possible. While the specification presented in this section
is the last version, which was implemented in the AHABus prototype, some parts
of the protocol evolved or had to be changed during the prototyping process.
\todo{Might not be necessary to say that here, given further sections}

@@

\todo{Add some details about simplicity of the network topology}

@@

\item[Frame Layer:] The frame layer relies on the Link Layer to transmit its
data units, and provides reliability and error correction to higher-level 
layers.\todo{Add vocabulary maybe (ancillary)}

@@

The design process and specification of each of those three layers is explained
in more details in the following sections.

\subsubsection{Link Layer}
\label{sssec:link-layer}

\todo{add an introductory sentence?}

@@

RTTY functions like UART, over radio. Instead of using logical electrical levels
(5V for logical high, 0V for logical low), it uses Audio Frequency Shift Keying
(\textit{AFSK}): two audio tones separated by a frequency shift act as two
logical levels. When nothing is being transmitted, the line is held high. The
start of a byte is signalled by the \textit{start bit} (the line goes low),
followed by the eight \textit{data bits}, and one or two \textit{stop bits} (the
line is held high). After a byte has been sent, the line is held high again
until the next byte, or the next byte starts immediately.

Other existing protocols exist that can transmit data over radio audio signals,
\todo{references for DominoEx and Olivia}
like DominoEx or Olivia, sometimes with better transmission rates. RTTY was
preferred over them because the ESP8266 board used as OBC in the AHABus
prototype provides hardware support for UART, making it easier to implement and
reliable. Since the link layer is transparent to the frame and packet layers,
RTTY could be swapped for another protocol in the future with no impact on
those.

The valid encoding and decoding of a RTTY signal depends on the chosen baud
rate (the number of times the logical signal changes level). During tests,
the highest baud rate that allowed clear decoding of radio signals at 10mW
power rating was 200bauds. Since each byte requires eleven bits to be sent
(one start bit, eight data bits and two stop bits), this corresponds to a
\(\frac{200}{11}=18.18\) bytes per second bandwidth.

@@

The role of the frame layer is to provide higher-level layers with a reliable
data stream: the Link Layer discussed in section \ref{sssec:link-layer} will
send data bytes one after another, but it does not guarantee that the bytes are
received, or that noise on the channel (in this case, radio waves) have not
damaged the data between transmission and reception. Those are the two main
potential types of issues that the Frame Layer was designed to address.

Preventing complete data loss on a radio channel is only possible to a degree:
the signal-to-noise ratio is limited by multiple factors including transmission
power – limited in the United Kingdom – distance, other neighbouring
transmissions and line-of-sight. Those issues could have been mitigated in part
through solutions that were outside of this project's scope, like antennae
design. While this meant that some data loss could not be prevented, a solution
was required to detect occurrences of data loss, and identify which part of the
data was missing.

Identifying lost sections of an arbitrary stream of bytes without any form of
formatting is not possible: since it is not known in advance what data should
be expected, there is no model to which actual received data can be compared.
For this reason, the AHABus radio protocol was designed as a framed protocol:
the stream of data is divided in chunks of a set number of bytes, encapsulated
in \textit{frames} that contain the a sequence number used to track the position
of the chunk in the data stream.

The AHABus frame design was inspired by other frame-based protocols like the
CCSDS Space Packet/Frame protocol. Each frame is composed of two section: a
header, that contains metadata, and a payload containing the data carried by
that frame.

The only piece of metadata strictly required to be carried in a frame was its
sequence number. The frame format was designed to use a 16-bit integer to
represent the sequence number, as it allowed a large number of frames to be
sent in-between wrap-arounds of the number (when it becomes larger than 65,535).

To ensure that frames can be identified in the future, even if the format were
to evolve, the protocol version number was also added to the header format. At
the time of this writing, the protocol version was 0. If a future frame decoder
were to receive a frame indicating an old protocol version, it could be at
least signalled to the user, if not handled gracefully.

\todo{explain requirement for sync marker}

The resulting first version of the frame format is presented in figure
\ref{fig:frame-fmt-orig}.

\begin{figure}[H]
    \begin{bytefield}{24}
        \bitheader{0,7,8,15,16,23} \\
        \bitbox{8}{version} & \bitbox{16}{sequence number} \\
            \wordbox[lrt]{1}{payload data} \\
            \skippedwords \\
            \wordbox[lrb]{1}{}
    \end{bytefield}
    \centering
    \caption{Original AHABus Frame Format}
    \label{fig:frame-fmt-orig}
\end{figure}

\todo{explain move away from double-marker towards sequence + mark}

Encapsulating the transmitted data adds overhead. If \(n\) is the size of a
chunk in bytes, and each frame adds \(m\) bytes of metadata, then only a
percentage \(\frac{n}{n+m}\) of the stream contains actual data. For reasons
explained further, the AHABus frames were chosen to be 256-byte long. As shown
previously, each frame also includes.

\begin{figure}[H]
    \begin{bytefield}{16}
        \bitheader{0,7,8,15} \\
        \begin{rightwordgroup}{Header}
            \bitbox{8}{01011010} & \bitbox{8}{version} \\
            \bitbox{16}{sequence number}
        \end{rightwordgroup} \\
        \begin{rightwordgroup}{Payload}
            \wordbox[lrt]{1}{data} \\
            \skippedwords \\
            \wordbox[lrb]{1}{}
        \end{rightwordgroup} \\
        \begin{rightwordgroup}{FEC}
            \wordbox[lrt]{1}{checksum} \\
            \skippedwords \\
            \wordbox[lrb]{1}{}
        \end{rightwordgroup}
    \end{bytefield}
    \centering
    \caption{AHABus Frame Format}
\end{figure}

% Topic list for this:
%   * Provide reliability:
%       * Data arrives whole
%       * We receive valid data
%   * receive all data: we chunk the data stream, number the chunks         v
%   * we tag data with housekeeping things                                  v~
%   * If frames too small, housekeeping/data ratio is shit
%   * Need a way to detect the start + end of a frame.
%   * Explain that end of frame marker complicated, going with count?
%
%   * Validity checked (and corrected using FEC)
%   * FEC: went with Reed-Solomon for simplicity + availability
%   * FEC means we need small blocks of data -> Frames
%
%   * FEC forces us to change from double marker inside to sync sequence + frame

@@

\begin{figure}[H]
    \begin{bytefield}{32}
        \bitheader{0,7,8,15,16,23,24,31} \\
        \begin{leftwordgroup}{Header}
            \bitbox{8}{version} & \bitbox{8}{payload ID} &
            \bitbox{16}{packet length} \\
            \bitbox{32}{latitude} \\
            \bitbox{32}{longitude} \\
            \bitbox{16}{altitude} & \bitbox[lrt]{16}{}
        \end{leftwordgroup} \\
        \begin{leftwordgroup}{Payload}
            \wordbox[lrt]{1}{data} \\
            \skippedwords \\
            \wordbox[lrb]{1}{}
        \end{leftwordgroup}
    \end{bytefield}
    \centering
    \caption{AHABus Frame Format}
\end{figure}

@@

\subsection{Architecture}

\subsection{User Interface}

@@

\subsection{Extracting Raw Binary Stream}

\subsection{Packet}

@@

\newpage
\chapter{Results}
\label{ch:results}
